Appendix

### 2023-10-30 manuscript/supplementaryRW.Rmd
corrections in suplementary doc

---
title: "Appendix"
subtitle: "Tree regeneration in models of forest dynamics: a key priority for further research"
author: 
- Olalla Díaz-Yáñez,
- Yannek Käber,
- Tim Anders,
- Friedrich Bohn,
- Kristin H. Braziunas,
- Josef	Brůna,
- Rico Fischer,
- Samuel M. Fischer,
- Jessica Hetzer,
- Thomas Hickler,
- Christian	Hochauer,
- Manfred J. Lexer,
- Heike	Lischke,
- Mats Mahnken,
- Paola	Mairota,
- Ján Merganič,
- Katarina Merganičová,
- Tobias Mette,
- Marco Mina,
- Xavier Morin,
- Werner Rammer,
- Christopher P.O. Reyer,
- Simon	Scheiter,
- Daniel Scherrer,
- Harald Bugmann
date: ""
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    latex_engine: xelatex
    extra_dependencies: ["flafter", "float", "subfig"]
    toc: yes
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: chicago-author-date.csl
editor_options: 
  markdown: 
    wrap: 72
fontsize: 12pt
linestretch: 2
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
- \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
- \usepackage{fontspec}
- \setmainfont{Times New Roman}
---

@@

# Appendix 1: Model reports

@@

Mahnken M., Gutsch M., Reyer C.P.O., Lasch-Born P.

@@

All species required were simulated: Fagus sylvatica, Picea abies, Abies
alba, Carpinus betulus, Tilia cordata, Acer pseudoplatanus, Betula spp.,
Fraxinus excelsior, Quercus spp., Alnus glutinosa, Pinus sylvestris

In addition the following species were simulated: Acer campestre, Acer
platanoides, Alnus incana, Alnus viridis, Castanea sativa, Corylus
avellana, Larix decidua, Pinus montana, Populus nigra, Populus tremula,
Salix alba, Sorbus aria, Sorbus aucuparia, Taxus baccata, Tilia
platyphyllos, Ulmus glabra, Pinus cembra

@@

Fabrika M., Merganič J., Merganičová K., 2009: Natural regeneration
density model developed for the purposes of the individual-tree growth
simulator. Acta Facultatis Forestalis Zvolen, 51(3), p. 125-137, ISSN:
0231-5785 Fabrika, M., (2005): Simulátor biodynamiky lesa SIBYLA,
koncepcia, konštrukcia a programové riešenie. Habilitačná práca.
Technická univerzita vo

Zvolene, 238 p. Kahn, M., (1994): Modellierung der Höhenentwicklung
ausgewählter Baumarten in Abhängigkeit vom Standort. Forstliche
Forschungsber. München, Vol. 141, 221 p. Keeling C. D., R. B. Bacastow,
A. E. Bainbridge, C. A. Ekdahl Jr., P. R. Guenther, L. S. Waterman, J.
F. S. Chin 1976: Atmospheric carbon dioxide variations at Mauna Loa
Observatory, Hawaii. Tellus. XXVIII, 6: 538-551,
<https://doi.org/10.1111/j.2153-3490.1976.tb00701.x>

@@

Pretzsch, H., (1992): Konzeption und Konstruktion von Wuchsmodellen für
Rein- und Mischbestände. Forstliche Forschungsberichte München, Nr.115,
358 p. Valent, P., Merganič J., Výbošťok J. 2018: Optimus -- Software
for Multi-criteria analysis. TU Zvolen.

@@

```{=tex}
\newpage
\beginsupplement
```
# Appendix 2: Supplementary figures and tables

@@

kable(zeroInflated,  format = "latex", caption = "Number of total samples, samples with zero recruitment and the proportion of zero recruitment per model and observed data.")|>
        kableExtra::kable_styling(latex_options = "HOLD_position")|>
        #kableExtra::landscape() |>
        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
```

@@

sigDiff <- readr::read_csv(here::here("figures", "sigH7_10.csv"))

kable(sigDiff,  format = "latex", caption = "Test of the significance of the differences of Shannon index in each model for 7 and 10 cm threshold.") |>
        kableExtra::kable_styling(latex_options = "HOLD_position")|>
        #kableExtra::landscape() |>
        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
```

```{r recrStand7, echo=FALSE, fig.cap="Mean Shannon index per site at recruitment and at the stand level both for observed and simulated values for the diameter threshold of 7cm. The black line marks where the mean Shannon index per site at recruitment and at the stand level would be equal", out.width='100%', fig.pos='H'}
knitr::include_graphics(here::here("figures", "recruitmentAdultRichness7.png"))
```

```{r recrStand10, echo=FALSE, fig.cap="Mean Shannon index per site at recruitment and at the stand level both for observed and simulated values for the diameter threshold of 10cm", out.width='100%', fig.pos='H'}
knitr::include_graphics(here::here("figures", "recruitmentAdultRichness10.png"))
```

```{r simObadult , echo=FALSE, fig.cap="Mean species richness at the stand level per site in the simulated versus observed species richness. The species richness is estimated using the Shannon index  calculated by basal area. A1) groups models where the simulated species composition in adults is higher than the observed, A2) groups models where the simulated species composition in adults is similar to the observed, A3) groups models where the simulated species composition in adults is lower than the observed. NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%', fig.pos='H'}
knitr::include_graphics(here::here("figures", "allAgesRichness.png"))
```

```{r simObr, echo=FALSE, fig.cap="Mean species richness per site in the simulated recruitment versus observed species richness in the recruitment. The species richness is estimated using the Shannon index using the basal area. B1) groups models where the simulated species composition in recruitment is higher than the observed, A2) groups models where the simulated species composition in recruitment is similar to the observed, A3) groups models where the simulated species composition in recruitment is lower than the observed. NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA", out.width='100%', fig.pos='H'}
knitr::include_graphics(here::here("figures", "recruitmentRichness.png"))
```

@@

kable(sigDiff,  format = "latex", caption = "Test of the significance of the differences of rate of recruitment at 7 cm threshold over 10 cm threshold between the observed values and the simulated values per model.") |>
    kableExtra::kable_styling(latex_options = "HOLD_position")|>
#kableExtra::landscape()
kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
```

@@

tabletrends <- readr::read_csv(here::here("figures", "mort7_10_trends.csv"))
kable(tabletrends,  format = "latex", caption = "Models trends in mortality between 7 and 10cm, and the ingrowth at 7cm, based in the mean oer site .")|>
kableExtra::kable_styling(latex_options = "HOLD_position")
```

```{r mortality3 , echo=FALSE, fig.cap="Recruitment ratio between 7 and 10cm and the initial recruitment tree number at 7cm. Dashed red lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke self-thinning ratio under evenaged conditions", out.width='70%', fig.pos='H'}
knitr::include_graphics(here::here("figures", "mort7_10_trends.png"))
```

@@

```{r plotComp, echo=FALSE, warning=FALSE, fig.cap="Plot and summary statistics of the linear model between the overestimation proportion in ingrowth levels and the mean complexity value of the establisment modules", fig.show = 'hold', out.width = '70%', fig.pos='H'}

@@

kable(rtreesDiff,  format = "latex", caption = "Test of the significance of the differences of rate of recruitment at 7 cm threshold of the observed values and the simulated values per model.") |>
    kableExtra::kable_styling(latex_options = "HOLD_position") |>
    kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
```

>>>>

---
title: "Appendix"
subtitle: "Tree regeneration in models of forest dynamics: a key priority for further research  ECOSPHERE"
author: 
- Olalla Díaz-Yáñez,
- Yannek Käber,
- Tim Anders,
- Friedrich Bohn,
- Kristin H. Braziunas,
- Josef	Brůna,
- Rico Fischer,
- Samuel M. Fischer,
- Jessica Hetzer,
- Thomas Hickler,
- Christian	Hochauer,
- Manfred J. Lexer,
- Heike	Lischke,
- Paola	Mairota,
- Ján Merganič,
- Katarina Merganičová,
- Tobias Mette,
- Marco Mina,
- Xavier Morin,
- Mats Nieberg,
- Werner Rammer,
- Christopher P.O. Reyer,
- Simon	Scheiter,
- Daniel Scherrer,
- Harald Bugmann
date: ""
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    latex_engine: xelatex
    extra_dependencies: ["flafter", "float", "subfig"]
    toc: yes
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: chicago-author-date.csl
editor_options: 
  markdown: 
    wrap: 72
fontsize: 12pt
linestretch: 2
header-includes:
- \usepackage[left]{lineno}
- \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
- \usepackage{fontspec}
- \setmainfont{Times New Roman}
---

@@

# Model reports

@@

Nieberg M., Gutsch M., Reyer C.P.O., Lasch-Born P.

@@

All species required were simulated: *Fagus sylvatica, Picea abies,
Abies alba, Carpinus betulus, Tilia cordata, Acer pseudoplatanus, Betula
spp., Fraxinus excelsior, Quercus spp., Alnus glutinosa, Pinus
sylvestris.*

In addition the following species were simulated: *Acer campestre, Acer
platanoides, Alnus incana, Alnus viridis, Castanea sativa, Corylus
avellana, Larix decidua, Pinus montana, Populus nigra, Populus tremula,
Salix alba, Sorbus aria, Sorbus aucuparia, Taxus baccata, Tilia
platyphyllos, Ulmus glabra, Pinus cembra*

@@

Fabrika M., Merganič J., Merganičová K., 2009: Natural regeneration
density model developed for the purposes of the individual-tree growth
simulator. Acta Facultatis Forestalis Zvolen, 51(3), p. 125-137, ISSN:
0231-5785

Fabrika, M., (2005): Simulátor biodynamiky lesa SIBYLA, koncepcia,
konštrukcia a programové riešenie. Habilitačná práca. Technická
univerzita vo Zvolene, 238 p.

Kahn, M., (1994): Modellierung der Höhenentwicklung ausgewählter
Baumarten in Abhängigkeit vom Standort. Forstliche Forschungsber.
München, Vol. 141, 221 p.

Keeling C. D., R. B. Bacastow, A. E. Bainbridge, C. A. Ekdahl Jr., P. R.
Guenther, L. S. Waterman, J. F. S. Chin 1976: Atmospheric carbon dioxide
variations at Mauna Loa Observatory, Hawaii. Tellus. XXVIII, 6: 538-551,
<https://doi.org/10.1111/j.2153-3490.1976.tb00701.x>

@@

Pretzsch, H., (1992): Konzeption und Konstruktion von Wuchsmodellen für
Rein- und Mischbestände. Forstliche Forschungsberichte München, Nr.115,
358 p.

Valent, P., Merganič J., Výbošťok J. 2018: Optimus -- Software for
Multi-criteria analysis. TU Zvolen.

@@

```{=tex}
\newpage
\beginsupplement
```
# Supplementary figures and tables

@@

kable(zeroInflated,  format = "latex", caption = "Number of total samples, samples with zero recruitment and the proportion of zero recruitment per model and observed data.") |>
        kableExtra::kable_styling(latex_options = "HOLD_position") |>
        kableExtra::kable_styling(full_width = F, latex_options = "scale_down")
```

@@

sigDiff <- readr::read_csv(here::here("figures", "sigH7_10.csv"))
sigDiff$p.adj <- NULL
colnames(sigDiff) <- c("model", "y", "Group 1", "Group 2", "p", "p.signif")

kable(sigDiff,  format = "latex", caption = "Test of the significance of the differences of Shannon index in each model for 7 and 10 cm threshold.") |>
        kableExtra::kable_styling(latex_options = "HOLD_position")
```

```{r recrStand7, echo=FALSE, fig.cap="Mean Shannon index per site at recruitment and at the stand level both for observed and simulated values for the diameter threshold of 7cm. The black line marks where the mean Shannon index per site at recruitment and at the stand level would be equal.", out.width='100%', fig.pos='H'}
knitr::include_graphics(here::here("figures", "recruitmentAdultRichness7.png"))
```

```{r recrStand10, echo=FALSE, fig.cap="Mean Shannon index per site at recruitment and at the stand level both for observed and simulated values for the diameter threshold of 10cm.", out.width='100%', fig.pos='H'}
knitr::include_graphics(here::here("figures", "recruitmentAdultRichness10.png"))
```

```{r simObadult , echo=FALSE, fig.cap="Mean species richness at the stand level per site in the simulated versus observed species richness. The species richness is estimated using the Shannon index  calculated by basal area. A1) groups models where the simulated species composition in adults is higher than the observed, A2) groups models where the simulated species composition in adults is similar to the observed, A3) groups models where the simulated species composition in adults is lower than the observed. NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA.", out.width='100%', fig.pos='H'}
knitr::include_graphics(here::here("figures", "allAgesRichness.png"))
```

```{r simObr, echo=FALSE, fig.cap="Mean species richness per site in the simulated recruitment versus observed species richness in the recruitment. The species richness is estimated using the Shannon index using the basal area. B1) groups models where the simulated species composition in recruitment is higher than the observed, A2) groups models where the simulated species composition in recruitment is similar to the observed, A3) groups models where the simulated species composition in recruitment is lower than the observed. NOTE: The recruitment treshold recruitment for treshold 7 cm in the empirical data has several sites with no recruitment or no data and the Shannon index for those sites is NA.", out.width='100%', fig.pos='H'}
knitr::include_graphics(here::here("figures", "recruitmentRichness.png"))
```

@@

kable(sigDiff,  format = "latex", caption = "Test of the significance of the differences of rate of recruitment at 7 cm threshold over 10 cm threshold between the observed values and the simulated values per model.") |>
    kableExtra::kable_styling(latex_options = "HOLD_position") 
```

@@

tabletrends <- readr::read_csv(here::here("figures", "mort7_10_trends.csv"))
tabletrends$Significance <- round(tabletrends$Significance, 4)
tabletrends$Slope <- round(tabletrends$Slope, 4)
kable(tabletrends,  format = "latex", caption = "Models trends in mortality between 7 and 10cm (represented in Figure S5), and the ingrowth at 7cm, based in the mean per site. The slope is the rate of change from the linear model between recruitment ratio 7-10 cm and the initial recruitment tree number.") |>
kableExtra::kable_styling(latex_options = "HOLD_position")
```

```{r mortality3 , echo=FALSE, fig.cap="Recruitment ratio between 7 and 10cm and the initial recruitment tree number at 7cm. Dashed red lines mark a ratio equal to 1 indicating no decrease between 7 and 10cm tree recruitment and ratio equal to 1.77 corresponding with the Reineke self-thinning ratio under evenaged conditions.", out.width='70%', fig.pos='H'}
knitr::include_graphics(here::here("figures", "mort7_10_trends.png"))
```

@@

```{r plotComp, echo=FALSE, warning=FALSE, fig.cap="Plot and summary statistics of the linear model between the overestimation proportion in ingrowth levels and the mean complexity value of the establisment modules.", fig.show = 'hold', out.width = '70%', fig.pos='H'}

@@

kable(rtreesDiff,  format = "latex", caption = "Test of the significance of the differences of rate of recruitment at 7 cm threshold of the observed values and the simulated values per model.") |>
    kableExtra::kable_styling(latex_options = "HOLD_position") 
```
